### 2024-07-20 mythbusters.qmd
20240719 - remove `here` package dependency
20240719 - contract year

```{r}
library("here")
library("tidyverse")
```

@@

```{r}
load(file = here("data", "nfl_playerContracts.RData"))
load(file = here("data", "nfl_pbp.RData"))
```

```{r}
#| eval: false
#| include: false

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
```

## Do Players Score More Fantasy Points in their Contract Year? {#sec-contractYear}

```{r}
# Subset to remove players without a year signed
nfl_playerContracts_subset <- nfl_playerContracts %>% 
  dplyr::filter(!is.na(year_signed) & year_signed != 0)

@@

# Check individual players
nfl_playerContracts_subset %>% 
  dplyr::filter(player == "Aaron Rodgers") %>% 
  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)

nfl_playerContracts_subset %>% 
  dplyr::filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)

@@

>>>>

```{r}
library("petersenlab")
library("lme4")
library("lmerTest")
library("tidyverse")
```

@@

```{r}
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = "./data/nfl_espnQBR.RData")
```

## Do Players Perform Better in their Contract Year? {#sec-contractYear}

```{r}
#| code-fold: true

# Subset to remove players without a year signed
nfl_playerContracts_subset <- nfl_playerContracts %>% 
  dplyr::filter(!is.na(year_signed) & year_signed != 0)

@@

# Check individual players
#nfl_playerContracts_subset %>% 
#  dplyr::filter(player == "Aaron Rodgers") %>% 
#  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)
#
#nfl_playerContracts_subset %>% 
#  dplyr::filter(player %in% c("Jared Allen", "Aaron Rodgers")) %>% 
#  dplyr::select(player:years, contractYear, next_contract_start, played_in_contract_year)

# Subset data
nfl_playerContractYears <- nfl_playerContracts_subset %>% 
  dplyr::filter(played_in_contract_year == TRUE) %>% 
  dplyr::filter(position %in% c("QB","RB","WR","TE")) %>% 
  dplyr::select(player, position, team, contractYear) %>% 
  dplyr::mutate(playerMerge = toupper(str_replace_all(player, "[[:punct:]]", ""))) %>% 
  dplyr::rename(season = contractYear) %>% 
  dplyr::mutate(contractYear = 1)

# Merge with weekly and seasonal stats data
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
  dplyr::mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
#nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
#  mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))

nfl_actualStatsContracts_offense_weekly <- dplyr::full_join(
  nfl_actualStats_offense_weekly,
  nfl_playerContractYears,
  by = c("playerMerge", "position_group" = "position", "season")
) %>% 
  dplyr::filter(position_group %in% c("QB","RB","WR","TE"))

#nfl_actualStatsContracts_offense_seasonal <- full_join(
#  nfl_actualStats_offense_seasonal,
#  nfl_playerContractYears,
#  by = c("playerMerge", "position_group" = "position", "season")
#) %>% 
#  filter(position_group %in% c("QB","RB","WR","TE"))

nfl_actualStatsContracts_offense_weekly$contractYear[which(is.na(nfl_actualStatsContracts_offense_weekly$contractYear))] <- 0
#nfl_actualStatsContracts_offense_seasonal$contractYear[which(is.na(nfl_actualStatsContracts_offense_seasonal$contractYear))] <- 0

nfl_actualStatsContracts_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
  dplyr::arrange(playerMerge, season, season_type, week)

#nfl_actualStatsContracts_offense_seasonal <- nfl_actualStatsContracts_offense_seasonal %>% 
#  arrange(playerMerge, season)

nfl_actualStatsContractsSubset_offense_weekly <- nfl_actualStatsContracts_offense_weekly %>% 
  dplyr::filter(season_type == "REG")

# Calculate Per Game Totals
nfl_actualStatsContracts_seasonal <- nfl_actualStatsContractsSubset_offense_weekly %>% 
  dplyr::group_by(playerMerge, season) %>% 
  dplyr::summarise(
    position_group = petersenlab::Mode(position_group),
    rushing_yards = sum(rushing_yards, na.rm = TRUE), # season total
    carries = sum(carries, na.rm = TRUE), # season total
    rushing_epa = mean(rushing_epa, na.rm = TRUE),
    receiving_yards = mean(receiving_yards, na.rm = TRUE),
    receiving_epa = mean(receiving_epa, na.rm = TRUE),
    contractYear = mean(contractYear, na.rm = TRUE),
    games = n(),
    .groups = "drop_last"
  ) %>% 
  dplyr::mutate(ypc = rushing_yards / carries)

@@

### QB

First, we prepare the data by merging and performing additional processing:

```{r}
#| code-fold: true

# Merge with QBR data
nfl_espnQBR$playerMerge <- paste(nfl_espnQBR$name_first, nfl_espnQBR$name_last, sep = " ") %>% 
  stringr::str_replace_all(., "[[:punct:]]", "") %>% 
  toupper()

nfl_contractYearQBR_weekly <- nfl_playerContractYears %>% 
  dplyr::filter(position == "QB") %>% 
  dplyr::full_join(
    .,
    nfl_espnQBR,
    by = c("playerMerge","team","season")
  )

nfl_contractYearQBR_weekly$contractYear[which(is.na(nfl_contractYearQBR_weekly$contractYear))] <- 0

nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::arrange(playerMerge, season, season_type, game_week)

nfl_contractYearQBRsubset_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::filter(
    season_type == "Regular",
    game_week != "Season Total") %>% 
  dplyr::mutate(game_week = as.numeric(game_week)) %>% 
  dplyr::arrange(playerMerge, season, season_type, game_week)

#hist(nfl_contractYearQBRsubset_weekly$qb_plays) # players have at least 20 dropbacks per game

# Calculate Per Game Totals
nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
  dplyr::group_by(playerMerge, season) %>% 
  dplyr::summarise(
    qbr = mean(qbr_total, na.rm = TRUE),
    pts_added = mean(pts_added, na.rm = TRUE),
    epa_pass = mean(pass, na.rm = TRUE),
    qb_plays = sum(qb_plays, na.rm = TRUE), # season total
    contractYear = mean(contractYear, na.rm = TRUE),
    games = n(),
    .groups = "drop_last"
  )

nfl_contractYearQBRsubset_seasonal <- nfl_contractYearQBR_seasonal %>% 
  dplyr::filter(games >= 5) # keep only player-season combinations in which QBs played at least 5 games
```

Then, we analyze the data.
Below is a mixed model that examines whether a player has a higher QBR per game when they are in a contract year compared to when they are not in a contract year.

```{r}
mixedModel_qbr <- lmerTest::lmer(
  qbr ~ contractYear + (1 | playerMerge),
  data = nfl_contractYearQBR_seasonal
)

summary(mixedModel_qbr)
```

```{r}
mixedModel_ptsAdded <- lmerTest::lmer(
  pts_added ~ contractYear + (1 | playerMerge),
  data = nfl_contractYearQBR_seasonal
)

summary(mixedModel_ptsAdded)
```

```{r}
mixedModel_epaPass <- lmerTest::lmer(
  epa_pass ~ contractYear + (1 | playerMerge),
  data = nfl_contractYearQBR_seasonal
)

summary(mixedModel_epaPass)
```

```{r}
# Placeholder for model predicting fantasy points
```

### RB

```{r}
#| code-fold: true

nfl_actualStatsContractsRB_seasonal <- nfl_actualStatsContracts_seasonal %>% 
  dplyr::filter(
    position_group == "RB",
    games >= 5)
```

```{r}
mixedModel_ypc <- lmerTest::lmer(
  ypc ~ contractYear + (1 | playerMerge),
  data = nfl_actualStatsContractsRB_seasonal
)

summary(mixedModel_ypc)
```

```{r}
mixedModel_epaRush <- lmerTest::lmer(
  rushing_epa ~ contractYear + (1 | playerMerge),
  data = nfl_actualStatsContractsRB_seasonal
)

summary(mixedModel_epaRush)
```

```{r}
# Placeholder for model predicting fantasy points
```

### WR/TE

```{r}
#| code-fold: true

nfl_actualStatsContractsWRTE_seasonal <- nfl_actualStatsContracts_seasonal %>% 
  dplyr::filter(
    position_group %in% c("WR","TE"),
    games >= 5)
```

```{r}
mixedModel_receivingYards <- lmerTest::lmer(
  receiving_yards ~ contractYear + (1 | playerMerge),
  data = nfl_actualStatsContractsWRTE_seasonal
)

summary(mixedModel_receivingYards)
```

```{r}
mixedModel_epaReceiving <- lmerTest::lmer(
  receiving_epa ~ contractYear + (1 | playerMerge),
  data = nfl_actualStatsContractsWRTE_seasonal
)

summary(mixedModel_epaReceiving)
```

```{r}
# Placeholder for model predicting fantasy points
```

### QB/RB/WR/TE

```{r}
# Placeholder for model predicting fantasy points
# Include player position as a covariate
```
